humanCdGF_predict_cfu

### 2016-10-22 code/HumanCdGF_predict_cfu.Rmd
add rf trained with one mouse per cage

# read in files
meta_data   <- read.table(meta_data, sep = '\t', header = T)
shared_file <- read.table(shared_file, sep = '\t', header = T, row.names = 'sample_id')
tax_file <- read.table(tax_file, sep = '\t', header = T, row.names = 'OTU')
source(tax_function)

@@

seed <- 1
n_trees <- 1001
rf_plot_list <- list()
n_features <- 10
cutoff <- 30

@@

ggplot(data = rf_prediction, aes(x=observed, y=predicted, color = cage_id)) + geom_point() + 
  labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)',
       title = "Random Forest Prediction of Colonized C Difficile CFU from Day 0 Community") + 
  theme_bw() + xlim(0,9) + ylim(0,9) + 
  annotate('text', x=0.15*(max(rf_prediction$observed)), 
           y=0.95*(max(rf_prediction$predicted)), parse = T,
           label=paste0('R^2 == ', round(tail(rf_n_features$rsq, 1), 2))) + 
  theme(legend.position = 'none')

>>>>

# read in files
meta_data   <- read.table(meta_data, sep = '\t', header = T, stringsAsFactors = F)
shared_file <- read.table(shared_file, sep = '\t', header = T, row.names = 'sample_id')
tax_file <- read.table(tax_file, sep = '\t', header = T, row.names = 'OTU')
source(tax_function)

@@

seed <- 1
n_trees <- 1001
n_features <- 10
cutoff <- 30

@@

##### - leave one cage out cv
LOcO_df <- data.frame(colonized_cfu %>% 
    right_join(community_d0.df) %>% 
    filter(!is.na(log_cfu_med)) %>% 
    select(log_cfu_med, mouse_id, cage_id, one_of(top_important_OTU)))
cage_list <- unique(LOcO_df$cage_id)

LOcO_probs <- c()
rsq <- c()
import_otus <- data.frame(otu = top_important_OTU)
for(i in 1:length(cage_list)){
  by_cage <- LOcO_df$cage_id == cage_list[i]
  LOcO_predict <- select(LOcO_df, -cage_id, -mouse_id)
  train <- LOcO_predict[!by_cage,]
  test <- LOcO_predict[by_cage,]
  set.seed(seed)
  temp_model <- randomForest(log_cfu_med~., data=train, pdel=0.99, ntree=500, importance=T)
  import_df <- importance(temp_model)
  colnames(import_df)[1] <- cage_list[i]
  import_otus <- merge(import_otus, import_df, by.x = 'otu', by.y = 'row.names')
  import_otus <- select(import_otus, -IncNodePurity)
  rsq[i] <- round(tail(temp_model$rsq, 1), 2)
  LOcO_probs <- rbind(LOcO_probs, data.frame(observed=test$log_cfu_med, 
                                             predicted=predict(temp_model, test),
                                             mouse_id=LOcO_df$mouse_id[by_cage],
                                             cage_id=rep(cage_list[i], nrow(test))))
}
med_otu_importance <- data.frame(otu = import_otus$otu,
                                 importance = apply(import_otus[, -1], 1, median))
med_otu_importance <- med_otu_importance[order(med_otu_importance$importance, decreasing = T),]

LOcO_rsq <- 1 - (sum((LOcO_probs$observed - LOcO_probs$predicted)^2)/
                   sum((LOcO_probs$observed - mean(LOcO_probs$observed))^2))

#One per cage
one_mouse_df <- data.frame(colonized_cfu %>% 
    right_join(community_d0.df) %>% 
    filter(!is.na(log_cfu_med))  %>% 
    select(log_cfu_med, mouse_id, cage_id, one_of(top_important_OTU)))

one_mouse_probs <- c()
om_rsq <- c()
om_rsq_rf <- c()
om_import_otus <- data.frame(otu = colnames(select(one_mouse_df, contains('Otu'))))
one_df <- c()
om_pred <- c()
for(i in 1:20){
    set.seed(i)
    train <- one_mouse_df %>% 
      group_by(cage_id) %>% 
      sample_n(1) %>% ungroup() 
    test <- one_mouse_df %>% 
      filter(!mouse_id %in% train$mouse_id)
    set.seed(seed)
    rf_om <- randomForest(log_cfu_med ~ ., data = select(train, -mouse_id, -cage_id),
                             importance = TRUE, ntree = n_trees)
    om_rsq_rf[i] <- tail(rf_om$rsq, 1)
    om_pred <- rbind(om_pred,
                     data.frame(select(test, mouse_id, cage_id, log_cfu_med),
                          predicted=predict(rf_om, test)))
    om_rsq[i] <- 1 - (sum((om_pred$log_cfu_med - om_pred$predicted)^2)/
               sum((om_pred$log_cfu_med - mean(om_pred$log_cfu_med))^2))
    import_df <- importance(rf_om)
    colnames(import_df)[1] <- paste(train$mouse_id, collapse = ' ')
    om_import_otus <- merge(om_import_otus, import_df, by.x = 'otu', by.y = 'row.names')
    om_import_otus <- select(om_import_otus, -IncNodePurity)
}

om_pred <- rename(om_pred, observed = log_cfu_med)

med_otu_importance_om <- data.frame(otu = om_import_otus$otu,
                                 importance = apply(import_otus[, -1], 1, median))
med_otu_importance_om <- med_otu_importance_om[order(med_otu_importance_om$importance, decreasing = T),]

#### merge OOB and leave 1 cage out dataframes
all_rf <- rbind(cbind(rf_prediction, data = rep('OOB', nrow(rf_prediction))),
                cbind(LOcO_probs, data = rep('LOO', nrow(LOcO_probs))),
                cbind(om_pred, data = rep('1_mouse', nrow(om_pred))))

ggplot(data = all_rf, aes(x=observed, y=predicted, color = data)) + geom_point() + 
  geom_abline(slope = 1) + 
  labs(x= 'Observed Log10 CFU', y = 'Predicted Log10 CFU',
       title = "Random Forest Prediction of Colonized C Difficile CFU from Day 0 Community") + 
  theme_bw() + annotate('text', x=1.05*(min(LOcO_probs$observed)), 
                        y=0.95*(max(LOcO_probs$predicted)), # parse = T,
                        label=paste0('OOB ', 'R^2 == ', round(tail(rf_n_features$rsq, 1), 2),
                                     '\nLOcO ','R^2 == ', round(LOcO_rsq, 2),
                                     '\n1 mouse', 'R^2 == ', round(tail(om_rsq, 1), 2))) + 
  theme(legend.position = 'none')

get_tax(1, med_otu_importance$otu, tax_file)
##### - leave one cage out cv

ggplot(data = rf_prediction, aes(x=observed, y=predicted, color = cage_id)) + geom_point() + 
  labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)',
       title = "Random Forest Prediction of Colonized C Difficile CFU from Day 0 Community") + 
  theme_bw() + annotate('text', x=1.05*(min(rf_prediction$observed)), 
                        y=0.95*(max(rf_prediction$predicted)), parse = T,
                        label=paste0('R^2 == ', round(tail(rf_n_features$rsq, 1), 2))) + 
  theme(legend.position = 'none')

importance.df <- data.frame(get_tax(1, names(importance_final), tax_file),
                            importance = importance_final)
tax_label <- paste0(importance.df$tax_label, ' : ', 
                    round(importance.df$importance, 3))
